---
title: "lab9_code"
format: html
editor: visual
---

## Lab 9: Generative Art

### Author: Rebecca Ioffe

```{r}

#| label: libraries

library(ggplot2)
library(tidyverse)
library(ggthemes)
library(ambient)
```

## Art Piece 1

#### Ordinary Plotting functions on data

Description of code choices:

```{r}
#clean dataset for segment plotting

iris <- iris %>%
  mutate(
    x0 = Sepal.Length,
    y0 = Sepal.Width,
    x1 = Petal.Length,
    y1 = Petal.Width,
    shade = Species,
    size = Petal.Width
  )

# Create custom color palate for species color segments

custom_colors <- c("#FF69B4", "#FF1493", "#DB7093")

# Create the segment plot
ggplot(iris, aes(
  x = x0,
  y = y0,
  xend = x1,
  yend = y1,
  color = shade,
  size = size
)) +
  geom_segment(show.legend = FALSE) +
  geom_point(aes(x = x1, y = y1), size = 3, shape = 18, color = "yellow", alpha = 0.7) +  
  # Add diamond-shaped, yellow, transparent points
  coord_polar() +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) + 
  scale_color_manual(values = custom_colors) + 
  theme_void() + 
  theme(plot.caption = element_text(hjust = 0)) +
  theme(
    plot.background = element_rect(fill = "black"),
    panel.background = element_rect(fill = "black"),
    plot.title = element_text(color = "white", hjust = 0.5),
    plot.subtitle = element_text(color = "white", hjust = 0.5),
    plot.caption = element_text(color = "white", hjust = 0.5)
  ) +
  labs(
    title = "Flower Wave",
    caption = "Polar Plot of Iris Dataset with Segments"
  )



```

Description:

The wave of bright colors, each representing a species of iris contrasts the dark background. Overlaid like fireflies are the end points of each segment: culminating at the edge of the wave. 


## Art Piece 2

#### Purely generative art piece

Description of code choices: 

```{r}

sample_canva <- function(seed = NULL) {
  if(!is.null(seed)) set.seed(seed)
  sample(ggthemes::canva_palettes, 1)[[1]]
}
```


```{r}
grid <- long_grid(seq(1, 10, length.out = 1000), seq(1, 10, length.out = 1000))


grid  %>%
  mutate(
    noise = gen_perlin(x,y, frequency = 0.9), 
    ripple = gen_waves(x,y, frequency = 5), 
    gate = gen_spheres(x,y, frequency = 25) %>% normalize(), 
    paint = noise + (1 + ripple) * (gate >= 0.1 & gate < 0.5)
  ) %>%
    ggplot(aes(x, y, fill = paint)) + 
    geom_raster(show.legend = FALSE) +
    theme_void() +
    coord_equal() +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_fill_gradientn(colours = sample_canva(seed = 54)) +
labs(
    title = "Functional Anatomy",
  )


```

Description: 

Inspired by the human body: red blood cells and veins flow underneath the surface of the skin. The concentric spheres represent the cyclical nature of human life with oblong ripples representing cells. The bright red and blue color palate pay tribute to the functions of the human body. 



